---
title: "Statistical Report"
author: "Otto"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(fpp3, skimr, tsibble, dplyr, readr, patchwork) 
BFR <- read_csv("BirthsAndFertilityRatesAnnual.csv")
```

# Statistical Report on Fertility Data in Singapore

```{r}
BFR <- BFR |>   
  filter(DataSeries %in% c("Total Fertility Rate (TFR)", "Total Live-Births")) |> 
  mutate(DataSeries = 
  recode(DataSeries,
         "Total Fertility Rate (TFR)" = "TFR",
         "Total Live-Births" = "TLB"  )) |>   
  select(DataSeries, `1960`:`2024`) |>   
  mutate(across(`1960`:`2024`, as.numeric)) |>   
  pivot_longer(     
    cols = `1960`:`2024`,     
    names_to = "Year",     
    values_to = "Value") |>   
  mutate(Year = as.integer(Year)) |>   
  pivot_wider(     
    names_from = DataSeries,     
    values_from = Value   ) |>   
  as_tsibble(index = Year)

BFR <- BFR |>
  mutate(
    d_log_TFR = difference(log(TFR), differences = 1),
    d_log_TLB = difference(log(TLB), differences = 1)
  )
```

## Abstract

## Introduction

**Can time series models trained on data from 1960–2012 accurately
forecast Singapore's continuing fertility decline over 2013–2024?**

TO INLCUDE WHEN INTRODUCTION IS DONE

Inlude nessacary abbreviations when writing it up

4.  Justify SARIMA period 12 in Introduction (page 3) The periodogram
    doesn't obviously point to period 12, so you need literature to
    justify it. Add a paragraph in the Introduction discussing the
    Chinese zodiac Dragon year effect on Singapore births. This is what
    makes the SARIMA(p,d,q)(P,D,Q)[12] models substantively motivated
    rather than just data-driven.

## Methods

The modelling workflow was designed to answer our forecasting needs. The
TFR and TLB series were extracted from the dataset, reshaped into annual
time series from 1960–2024 and split into a training period from
1960–2012 and a testing period from 2013–2024.

```{r}
train <- BFR |> filter(Year <= 2012)
test  <- BFR |> filter(Year >= 2013)
```

Time plots and the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) tests of the
two series were used to check for trends, changes in variance and the
transformations required before modelling. Both series showed evidence
of non-stationarity and since ARIMA models require the time series to be
stationary, transformation of the series was necessary prior to
modelling.

```{r tfr-plot, fig.cap="Annual Total Fertility Rate in Singapore"}
BFR |>
  autoplot(TFR) +
  labs(y = "Total Fertility Rate", x = "Year")
```

```{r tlb-plot, fig.cap="Total Live Births in Singapore"}
BFR |>
  autoplot(TLB)+
  labs(y= "Total Live Births", x = "Year")
```

The log transformation was used to stabilise variance and make the model
proportional rather than absolute changes. Differencing was used to
remove the trends shown inside the data.

Both series were then transformed using either just the log or the log
with first or second differencing. These transformed series were then
tested by the KPSS test to ensure an improvement of the stationarity of
the series.

```{r kpss, eval=FALSE, include=FALSE}
# KPSS on training data for the two transformations used in your models
train |> features(difference(log(TFR)), unitroot_kpss)
train |> features(difference(log(TLB)), unitroot_kpss)
train |> features(difference(log(TFR), differences = 2), unitroot_kpss)
train |> features(difference(log(TLB), differences = 2), unitroot_kpss)
```

Two transformations showed potential as being stationary
(first-differenced and logged, and second-differenced logged). With the
KPSS test a small p-value shows evidence against stationarity. For TFR,
the KPSS test gave that $p= 0.0406$ for the first-differnced logged
series and $p=0.1$ for the second-differenced logged series. For the
logged TLB with both first and second differencing gave the same
p-values of $p=0.1$. So for TLB stationarity was not rejected for either
transformation.

Despite the first difference of the TFR data having a KPSS test result
below the significance threshold, we would prefer there to be less
differencing as we do not want to over-difference the data and introduce
noise in a relatively short data-set. We kept the second-differenced
models as comparison as well so we could see another set of models in
the results.

Because the transformed series and baseline residual diagnostics showed
dependence around lags 11–13, higher-lag ARIMA models were also
considered. These models were included to capture longer-lag
autoregressive dependence that was not represented by the low-order
baseline models.

SARIMA-style models with period 12 were also considered. Although the
data are annual, the period-12 structure was motivated by both the
lag-12 dependence observed in the transformed series and the literature
discussed in the Introduction on Chinese zodiac-related birth timing
effects in Singapore. These models were included to assess whether a
12-year pattern could be represented more simply than by unrestricted
high-order ARIMA models. For the SARIMA candidates, the ACF and PACF
plots of the transformed series were inspected to guide the initial
period-12 model specification.

Models were only considered viable if their residuals were approximately
white noise. This again was assessed with ACF and PACF, portmanteau
tests and residual plots. Among viable models, the final model selection
was based on AIC and forecast accuracy over the 2013-2024 period.
Forecast accuracy was assessed with RMSE and MAE, with formulas provided
in the appendix.

- So informed decision to try ARMIA (12,x, y) and SARIMA models
  potentially

| Series | Models |
|------------------------------------|------------------------------------|
| TFR | SARIMA(0,1,1)(1,0,0)[12], SARIMA(1,1,1)(1,0,0)[12], SARIMA(1,2,1)(1,0,0)[12] |
| TLB | SARIMA(0,1,1)(1,0,0)[12], SARIMA(1,1,1)(1,0,0)[12], SARIMA(0,2,1)(1,0,0)[12] |

## Results

The ACF and PACF plots of the log-first-differenced series are shown in
Figure 3. For both TFR and TLB the plots show a remaining dependence on
lags 11-13, particularly at lag 12. This suggests that low order ARIMA
models may not fully account for the higher order dependencies shown by
Figure 3.

```{r acf-pacf-log-first-diff, fig.cap="ACF and PACF plots of the log-first-differenced TFR and TLB series"}
acf_tfrD <- train |>
  ACF(difference(log(TFR))) |>
  autoplot() +
  labs(title = "ACF of log-differenced TFR")

pacf_tfrD <- train |>
  PACF(difference(log(TFR))) |>
  autoplot() +
  labs(title = "PACF of log-differenced TFR")

acf_tlbD <- train |>
  ACF(difference(log(TLB))) |>
  autoplot() +
  labs(title = "ACF of log-differenced TLB")

pacf_tlbD <- train |>
  PACF(difference(log(TLB))) |>
  autoplot() +
  labs(title = "PACF of log-differenced TLB")

(acf_tfrD | pacf_tfrD) /
(acf_tlbD | pacf_tlbD)
```

```{r acf-pacf-log-second-diff, fig.cap="ACF and PACF plots of the log-second-differenced TFR and TLB series", include=FALSE}
acf_tfrDD <- train |>
  ACF(difference(log(TFR), differences = 2)) |>
  autoplot() +
  labs(title = "ACF of log-second-differenced TFR")

pacf_tfrDD <- train |>
  PACF(difference(log(TFR), differences = 2)) |>
  autoplot() +
  labs(title = "PACF of log-second-differenced TFR")

acf_tlbDD <- train |>
  ACF(difference(log(TLB), differences = 2)) |>
  autoplot() +
  labs(title = "ACF of log-second-differenced TLB")

pacf_tlbDD <- train |>
  PACF(difference(log(TLB), differences = 2)) |>
  autoplot() +
  labs(title = "PACF of log-second-differenced TLB")

(acf_tfrDD | pacf_tfrDD) /
(acf_tlbDD | pacf_tlbDD)
```


To assess the period-12 SARIMA candidates, the ACF and PACF of $(1-B)\log(X_{t})$ (Figure 3) and $(1-B^{12})(1-B)\log(X_{t})$ (Figure 4) were used to guide the structure of our models. For both series, Figure 3 showed PACF spikes at lag-12 this suggested a seasonal AR term could be useful, while the lower lag ACF structure supported testing simple MA terms. This informed our decision for our first SARIMA candidate model to be SARIMA(0,1,1)(1,0,0)[12]. Nearby candidates were also fitted by varying the non-seasonal AR and MA orders and by comparing seasonal AR and MA terms.

In our seasonally differenced plot (Figure 4), lag-12 dependence remained even after the period 12 differencing. This indicated that differencing alone could not remove the period 12 dependence structure and supported the use of seasonal ARIMA models. 

```{r acf-pacf-seasonal-diff, fig.cap="ACF and PACF plots of the log-first-differenced and period-12 differenced TFR and TLB series"}

acf_tfrSD <- train |>
  ACF(difference(difference(log(TFR)), lag = 12)) |>
  autoplot() +
  labs(title = "ACF of period-12 differenced log TFR")

pacf_tfrSD <- train |>
  PACF(difference(difference(log(TFR)), lag = 12)) |>
  autoplot() +
  labs(title = "PACF of period-12 differenced log TFR")

acf_tlbSD <- train |>
  ACF(difference(difference(log(TLB)), lag = 12)) |>
  autoplot() +
  labs(title = "ACF of period-12 differenced log TLB")

pacf_tlbSD <- train |>
  PACF(difference(difference(log(TLB)), lag = 12)) |>
  autoplot() +
  labs(title = "PACF of period-12 differenced log TLB")

(acf_tfrSD | pacf_tfrSD) /
(acf_tlbSD | pacf_tlbSD)
```

Firstly, we fitted some low-order ARIMA models without seasonal terms,
however, as expected based on the lag-12 dependence identified with the
ACF/PACF plots these models failed to produce white noise residuals. We
next decided to fit some higher-order ARIMA models with seasonal terms as well as some SARIMA models with lag 12.
You can see the tested models below with their resultant residuals:


| Model                      | of Notes    |
|:---------------------------|:---------|
| ARIMA(11,1,0)              |          |
| ARIMA(12,1,0)              |          |
| ARIMA(13,1,0)              |          |
| ARIMA(13,1,1)              |          |
| SARIMA(0,1,1)(1,0,0)[12]   |          |
| SARIMA(0,1,1)(0,0,1)[12]   |          |
| SARIMA(1,1,1)(1,0,0)[12]   |          |
| SARIMA(1,1,1)(0,0,1)[12]   |          |
| ARIMA(0,2,1)               |          |
| ARIMA(1,2,1)               |          |
| SARIMA(0,2,1)(1,0,0)[12]   |          |
| SARIMA(1,2,1)(1,0,0)[12]   |          |

```{r}
fit_tfr <- train |>
  model(
    # Higher-lag ARIMA models
    tfr_ar11_log = ARIMA(log(TFR) ~ 0 + pdq(11, 1, 0),
                         order_constraint = TRUE),

    tfr_ar12_log = ARIMA(log(TFR) ~ 0 + pdq(12, 1, 0),
                         order_constraint = TRUE),

    tfr_ar13_log = ARIMA(log(TFR) ~ 0 + pdq(13, 1, 0),
                         order_constraint = TRUE),

    tfr_ar13ma1_log = ARIMA(log(TFR) ~ 0 + pdq(13, 1, 1),
                            order_constraint = TRUE),

    # SARIMA-style period-12 models
    tfr_sar_011_100_log = ARIMA(log(TFR) ~ 0 + pdq(0, 1, 1) + 
                                  PDQ(1, 0, 0, period = 12),
                                order_constraint = TRUE),

    tfr_sar_011_001_log = ARIMA(log(TFR) ~ 0 + pdq(0, 1, 1) + 
                                  PDQ(0, 0, 1, period = 12),
                                order_constraint = TRUE),

    tfr_sar_111_100_log = ARIMA(log(TFR) ~ 0 + pdq(1, 1, 1) + 
                                  PDQ(1, 0, 0, period = 12),
                                order_constraint = TRUE),

    tfr_sar_111_001_log = ARIMA(log(TFR) ~ 0 + pdq(1, 1, 1) + 
                                  PDQ(0, 0, 1, period = 12),
                                order_constraint = TRUE),

    # d=2 models
    tfr_021_log = ARIMA(log(TFR) ~ 0 + pdq(0, 2, 1)),
    
    tfr_121_log = ARIMA(log(TFR) ~ 0 + pdq(1, 2, 1)),
    
    tfr_sar_021_100_log = ARIMA(log(TFR) ~ 0 + pdq(0, 2, 1) + 
                                  PDQ(1, 0, 0, period = 12)),

    tfr_sar_121_100_log = ARIMA(log(TFR) ~ 0 + pdq(1, 2, 1) + 
                                  PDQ(1, 0, 0, period = 12))
  )
fit_tlb <- train |>
  model(
    # Higher-lag ARIMA models
    tlb_ar11_log = ARIMA(log(TLB) ~ 0 + pdq(11, 1, 0),
                         order_constraint = TRUE),

    tlb_ar12_log = ARIMA(log(TLB) ~ 0 + pdq(12, 1, 0),
                         order_constraint = TRUE),

    tlb_ar13_log = ARIMA(log(TLB) ~ 0 + pdq(13, 1, 0),
                         order_constraint = TRUE),

    tlb_ar13ma1_log = ARIMA(log(TLB) ~ 0 + pdq(13, 1, 1),
                            order_constraint = TRUE),

    # SARIMA-style period-12 models
    tlb_sar_011_100_log = ARIMA(log(TLB) ~ 0 + pdq(0, 1, 1) + 
                                  PDQ(1, 0, 0, period = 12),
                                order_constraint = TRUE),

    tlb_sar_011_001_log = ARIMA(log(TLB) ~ 0 + pdq(0, 1, 1) + 
                                  PDQ(0, 0, 1, period = 12),
                                order_constraint = TRUE),

    tlb_sar_111_100_log = ARIMA(log(TLB) ~ 0 + pdq(1, 1, 1) + 
                                  PDQ(1, 0, 0, period = 12),
                                order_constraint = TRUE),

    tlb_sar_111_001_log = ARIMA(log(TLB) ~ 0 + pdq(1, 1, 1) + 
                                  PDQ(0, 0, 1, period = 12),
                                order_constraint = TRUE),

    # d=2 models
    
    tlb_021_log = ARIMA(log(TLB) ~ 0 + pdq(0, 2, 1)),
    
    tlb_121_log = ARIMA(log(TLB) ~ 0 + pdq(1, 2, 1)),
    
    tlb_sar_021_100_log = ARIMA(log(TLB) ~ 0 + pdq(0, 2, 1) + 
                                  PDQ(1, 0, 0, period = 12)),

    tlb_sar_121_100_log = ARIMA(log(TLB) ~ 0 + pdq(1, 2, 1) + 
                                  PDQ(1, 0, 0, period = 12))
  )
```

```{r}
tfr_ljung <- bind_rows(
  fit_tfr |> select(tfr_ar11_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 11),

  fit_tfr |> select(tfr_ar12_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 12),

  fit_tfr |> select(tfr_ar13_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 13),

  fit_tfr |> select(tfr_ar13ma1_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 14),

  fit_tfr |> select(tfr_sar_011_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tfr |> select(tfr_sar_011_001_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tfr |> select(tfr_sar_111_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3),

  fit_tfr |> select(tfr_sar_111_001_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3),

  fit_tfr |> select(tfr_021_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 1),

  fit_tfr |> select(tfr_121_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tfr |> select(tfr_sar_021_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tfr |> select(tfr_sar_121_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3)
)
```

```{r}
tlb_ljung <- bind_rows(
  fit_tlb |> select(tlb_ar11_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 11),

  fit_tlb |> select(tlb_ar12_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 12),

  fit_tlb |> select(tlb_ar13_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 13),

  fit_tlb |> select(tlb_ar13ma1_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 14),

  fit_tlb |> select(tlb_sar_011_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tlb |> select(tlb_sar_011_001_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tlb |> select(tlb_sar_111_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3),

  fit_tlb |> select(tlb_sar_111_001_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3),

  fit_tlb |> select(tlb_021_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 1),

  fit_tlb |> select(tlb_121_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tlb |> select(tlb_sar_021_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 2),

  fit_tlb |> select(tlb_sar_121_100_log) |> augment() |>
    features(.innov, ljung_box, lag = 24, dof = 3)
)
```
```{r}
tfr_ljung <- tfr_ljung |>
  mutate(
    viable = if_else(lb_pvalue > 0.05,
                     "Viable: residual autocorrelation not detected",
                     "Not viable: residual autocorrelation detected")
  )

tlb_ljung <- tlb_ljung |>
  mutate(
    viable = if_else(lb_pvalue > 0.05,
                     "Viable: residual autocorrelation not detected",
                     "Not viable: residual autocorrelation detected")
  )

knitr::kable(
  tfr_ljung,
  digits = 3,
  caption = "Ljung-Box residual tests for TFR candidate models"
)

knitr::kable(
  tlb_ljung,
  digits = 3,
  caption = "Ljung-Box residual tests for TLB candidate models"
)

```
We then tested each of these models

Your professor expects the Results section to show:

ACF/PACF plots of the detrended and deseasonalised series How the first
viable model was identified from those plots How you varied parameters
to get additional viable models How each viable model fits the testing
data

## Discussion

## Conclusion

## References

## Appendices if applicable.
