---
title: "Statistical Report"
author: "Otto"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(fpp3, skimr, tsibble, dplyr, readr) 
BFR <- read_csv("BirthsAndFertilityRatesAnnual.csv")
```

# Statistical Report on Fertility Data in Singapore

```{r}
BFR <- BFR |>   
  filter(DataSeries %in% c("Total Fertility Rate (TFR)", "Total Live-Births")) |> 
  mutate(DataSeries = 
  recode(DataSeries,
         "Total Fertility Rate (TFR)" = "TFR",
         "Total Live-Births" = "TLB"  )) |>   
  select(DataSeries, `1960`:`2024`) |>   
  mutate(across(`1960`:`2024`, as.numeric)) |>   
  pivot_longer(     
    cols = `1960`:`2024`,     
    names_to = "Year",     
    values_to = "Value") |>   
  mutate(Year = as.integer(Year)) |>   
  pivot_wider(     
    names_from = DataSeries,     
    values_from = Value   ) |>   
  as_tsibble(index = Year)

BFR <- BFR |>
  mutate(
    d_log_TFR = difference(log(TFR), differences = 1),
    d_log_TLB = difference(log(TLB), differences = 1)
  )
```

## Abstract

## Introduction

**Can time series models trained on data from 1960–2012 accurately
forecast Singapore's continuing fertility decline over 2013–2024?**

TO INLCUDE WHEN INTRODUCTION IS DONE

Inlude nessacary abbreviations when writing it up

4. Justify SARIMA period 12 in Introduction (page 3)
The periodogram doesn't obviously point to period 12, so you need literature to justify it. Add a paragraph in the Introduction discussing the Chinese zodiac Dragon year effect on Singapore births. This is what makes the SARIMA(p,d,q)(P,D,Q)[12] models substantively motivated rather than just data-driven.



## Methods

The modelling workflow was designed to answer our forecasting needs. The
TFR and TLB series were extracted from the dataset, reshaped into annual
time series from 1960–2024 and split into a training period from
1960–2012 and a testing period from 2013–2024.

```{r}
train <- BFR |> filter(Year <= 2012)
test  <- BFR |> filter(Year >= 2013)
```

Time plots and the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) tests of the two series were used to check for trends, changes in variance and the transformations required before modelling. Both series
showed evidence of non-stationarity and since ARIMA models require the
time series to be stationary, transformation of the series was necessary
prior to modelling.

```{r tfr-plot, fig.cap="Annual Total Fertility Rate in Singapore"}
BFR |>
  autoplot(TFR) +
  labs(y = "Total Fertility Rate", x = "Year")
```

```{r tlb-plot, fig.cap="Total Live Births in Singapore"}
BFR |>
  autoplot(TLB)+
  labs(y= "Total Live Births", x = "Year")
```
The log transformation was used to stabilise variance and make the model proportional rather than absolute changes. Differencing was used to remove the trends shown inside the data. 

Both series were then transformed using either just the log or the log
with first or second differencing. These transformed series were then
tested by the KPSS test to ensure an improvement of the stationarity of the series. 

```{r eval=FALSE, include=FALSE}
# KPSS on training data for the two transformations used in your models
train |> features(difference(log(TFR)), unitroot_kpss)
train |> features(difference(log(TLB)), unitroot_kpss)
train |> features(difference(log(TFR), differences = 2), unitroot_kpss)
train |> features(difference(log(TLB), differences = 2), unitroot_kpss)
```

Two transformations showed potential as being stationary
(first-differenced and logged, and second-differenced logged). With the KPSS test
a small p-value shows evidence against stationarity.  For TFR, the KPSS test gave
that $p= 0.0406$ for the first-differnced logged series and $p=0.1$ for the
second-differenced logged series. For the logged TLB with both first and second
differencing gave the same p-values of $p=0.1$. So for TLB stationarity was not
rejected for either transformation. 


Despite the first difference of the TFR data having a KPSS test result
below the significance threshold, we would prefer there to be less
differencing as we do not want to over-difference the data and introduce
noise in a relatively short data-set. We kept the second-differenced models as comparison as well so we could see another set of models in the results.

Because the transformed series and baseline residual diagnostics showed dependence around lags 11–13, higher-lag ARIMA models were also considered. These models were included to capture longer-lag autoregressive dependence that was not represented by the low-order baseline models.

SARIMA-style models with period 12 were also considered. Although the data are annual, the period-12 structure was motivated by both the lag-12 dependence observed in the transformed series and the literature discussed in the Introduction on Chinese zodiac-related birth timing effects in Singapore. These models were included to assess whether a 12-year pattern could be represented more simply than by unrestricted high-order ARIMA models. For the SARIMA candidates, the ACF and PACF plots of the transformed series were inspected to guide the initial period-12 model specification.

Models were only considered viable if their residuals were approximately white
noise. This again was assessed with ACF and PACF, portmanteau tests and residual
plots. Among viable models, the final model selection was based on AIC and
forecast accuracy over the 2013-2024 period. Forecast accuracy was assessed with
RMSE and MAE, with formulas provided in the appendix. 

  - So informed decision to try ARMIA (12,x, y) and SARIMA models
    potentially

| Series | Models |
|------------------------------------|------------------------------------|
| TFR | SARIMA(0,1,1)(1,0,0)[12], SARIMA(1,1,1)(1,0,0)[12], SARIMA(1,2,1)(1,0,0)[12] |
| TLB | SARIMA(0,1,1)(1,0,0)[12], SARIMA(1,1,1)(1,0,0)[12], SARIMA(0,2,1)(1,0,0)[12] |

## Results

Your professor expects the Results section to show:

ACF/PACF plots of the detrended and deseasonalised series
How the first viable model was identified from those plots
How you varied parameters to get additional viable models
How each viable model fits the testing data

## Discussion

## Conclusion

## References

## Appendices if applicable.
