---
title: "Statistical Report"
author: "Otto"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(fpp3, skimr, tsibble, dplyr, readr) 
BFR <- read_csv("BirthsAndFertilityRatesAnnual.csv")
```

# Statistical Report on Fertility Data in Singapore

```{r}
BFR <- BFR |>   
  filter(DataSeries %in% c("Total Fertility Rate (TFR)", "Total Live-Births")) |> 
  mutate(DataSeries = 
  recode(DataSeries,
         "Total Fertility Rate (TFR)" = "TFR",
         "Total Live-Births" = "TLB"  )) |>   
  select(DataSeries, `1960`:`2024`) |>   
  mutate(across(`1960`:`2024`, as.numeric)) |>   
  pivot_longer(     
    cols = `1960`:`2024`,     
    names_to = "Year",     
    values_to = "Value") |>   
  mutate(Year = as.integer(Year)) |>   
  pivot_wider(     
    names_from = DataSeries,     
    values_from = Value   ) |>   
  as_tsibble(index = Year)

BFR <- BFR |>
  mutate(
    d_log_TFR = difference(log(TFR), differences = 1),
    d_log_TLB = difference(log(TLB), differences = 1)
  )

train <- BFR |> filter(Year <= 2012)
test  <- BFR |> filter(Year >= 2013)
```

## Abstract

## Introduction

**Can time series models trained on data from 1960–2012 accurately
forecast Singapore's continuing fertility decline over 2013–2024?**

```{r tfr-plot, fig.cap="Total Fertility Rate in Singapore, 1960–2024", results='markup'}
BFR |>
  autoplot(TFR)
BFR |> 
  autoplot(d_log_TFR)

```

```{r}
BFR |>
  autoplot(TLB)
BFR |> 
  autoplot(d_log_TLB)
```

```{r}

```

## Methods

- What the data is

<!-- -->

- talk about how the model's were chosen

  - Looked at transformation to make the data stationary

  - Then chose to log diff as it had the best \_ stationarity i think

    - The log transformation addresses **variance non-stationarity**,
      and the differencing addresses **mean non-stationarity** (the
      trend). They solve different problems. The KPSS test would confirm
      mean-stationarity after differencing, not "best stationarity"
      between log-diff and plain-diff.

  - The data still had trend for yr=12 (zodiac i think)

  - So informed decision to try ARMIA (12,x, y) and SARIMA models
    potentially

## Results

## Discussion

## Conclusion

## References

## Appendices if applicable.
