---
title: "Methods"
author: "Otto"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(fpp3, skimr, tsibble, dplyr, readr) 
BFR <- read_csv("BirthsAndFertilityRatesAnnual.csv")
BFR <- BFR |>   
  filter(DataSeries %in% c("Total Fertility Rate (TFR)", "Total Live-Births")) |> 
  mutate(DataSeries = 
  recode(DataSeries,
         "Total Fertility Rate (TFR)" = "TFR",
         "Total Live-Births" = "TLB"  )) |>   
  select(DataSeries, `1960`:`2024`) |>   
  mutate(across(`1960`:`2024`, as.numeric)) |>   
  pivot_longer(     
    cols = `1960`:`2024`,     
    names_to = "Year",     
    values_to = "Value") |>   
  mutate(Year = as.integer(Year)) |>   
  pivot_wider(     
    names_from = DataSeries,     
    values_from = Value   ) |>   
  as_tsibble(index = Year)

BFR <- BFR |>
  mutate(
    d_log_TFR = difference(log(TFR), differences = 1),
    d_log_TLB = difference(log(TLB), differences = 1)
  )
```

## Methods

The modelling workflow was designed to answer our forecasting needs. The
TFR and TLB series were extracted from the dataset, reshaped into annual
time series from 1960–2024 and split into a training period from
1960–2012 and a testing period from 2013–2024.

```{r}
train <- BFR |> filter(Year <= 2012)
test  <- BFR |> filter(Year >= 2013)
```

Time plots, Kwiatkowski-Phillips-Schmidt-Shin (KPSS) tests and ACF/PACF
plots of the two series were used to check for trends, changes in
variance and overall appropriateness for ARIMA modelling. Both series
showed evidence of non-stationarity and since ARIMA models require the
time series to be stationary, transformation of the series was necessary
prior to modelling.

```{r tfr-plot, fig.cap="Annual Total Fertility Rate in Singapore"}
BFR |>
  autoplot(TFR) +
  labs(y = "Total Fertility Rate", x = "Year")
```

```{r tlb-plot, fig.cap="Total Live Births in Singapore"}
BFR |>
  autoplot(TLB)+
  labs(y= "Total Live Births", x = "Year")
```

Both series were then transformed using either just the log or the log
with first or second differencing. These transformed series were then
tested by the KPSS test, residual plots and looking at the ACF and PACF
plots of the series to ensure that the transformation improves the
stationarity.

The KPSS test is used for testing whether a given time series is
stationary (Should I go into more detail or is this the level or should
it be in the appendix if it is?). The time series' are logged to scale
the variance proportionally if there is a trend so to keep relative
magnitude of changes in the data in the mind of the model. Differencing
is used for the removal of the trends shown inside the data.

```{r eval=FALSE, include=FALSE}
# KPSS on training data for the two transformations used in your models
train |> features(difference(log(TFR)), unitroot_kpss)
train |> features(difference(log(TLB)), unitroot_kpss)
train |> features(difference(log(TFR), differences = 2), unitroot_kpss)
train |> features(difference(log(TLB), differences = 2), unitroot_kpss)
```

Two transformations showed potential as being stationary
(first-differenced and logged, and second differenced logged). The TFR
data's KPSS test gave two different p-values for the two differencing
settings $p_{log D=1} = 0.0406$ and $p_{log D=2} = 0.1$. The logged
first difference setting's p-value was below the $0.05$, however, only
slightly so it may still be considered if a potential model can produce
white-noise residuals. For the TLB data both p-values were over the
significance threshold with $p=0.1$ for both of them meaning that we can
retain the null hypothesis, such that the TLB data is stationary with
either of the transformations.

Despite the first difference of the TFR data having a KPSS test result
below the significance threshold, we would prefer there to be less
differencing as we do not want to over-difference the data and introduce
noise in a relatively short data-set.

The next step was to pick appropriate models for forecasting for the
testing period. ACF and PACF plots were used to see if the time series
had hidden correlation which had carried thorugh the transformations.
Each time series transformation showed some dependence around lags 11-13
with the most significant spike at lag 12. This potential seasonality
made us consider models which would account for this correlation. Models
were only considered viable only if their residuals were approximately
white noise. Among viable models, the final model selection was based on
AIC and forecast accuracy over the 2013-2024 period.
